On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2015-09-11 MS.tex
Editing changes

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Anais Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia\\
\textasteriskcentered Email for correspondence: \texttt{ anais.gibert@gmail.com}\\
Word count: ~XXXX words}
\date{}

@@

\item Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits using a meta-analysis of 112 studies ($>$ 500 correlations). As a first approximation of plants size, we analysed the shift in the correlations between growth and traits across three plant stages (i.e. juvenile, sapling and adult).

@@

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Wright:2004jb,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, a growing number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with high photosynthetic rates and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation is mostly much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The growing discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

@@

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between trait and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a first approximation of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

Low WD decreases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table \ref{tab:trade-offs}), which should result in faster growth. Although cheaper wood carries a risk of increased mortality, this cost is not reflected in growth outcomes. Thus decreasing WD is predicted to increase growth across the entire size range.

@@

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed in either absolute (AGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and either SLA or SM in seedlings revert if absolute AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). For SLA, the correlation with RGR$_\textrm{mass}$ is positive but becomes negative with AGR$_\textrm{mass}$. For SM the correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For the others traits (WD, Hmax and Aarea), the predicted relationships are not affected by the growth measurement used.

@@

We aimed to identify studies that measured the correlation between five prominent functional trait (SLA, WD, Aarea, Hmax and SM) and a growth measurement across species. Our literature search was conducted in two phases.

@@

While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to meta-data, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

@@

Overall, we recorded coefficient correlations for a list of 20 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 15 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. SM and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost (\citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 20 traits were captured by a search on SLA and WD.

@@

We were unable to obtain a consistent and continuous measure of plant size across all studies, thus we relied on categorical groupings by ``stage'' to test our hypotheses. The authors of the 112 articles used in the meta-analyses, reported their plants as belonging to any of fix different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, by compiling data across multiple studies and vegetation types, we were concerned about the risk of assigning plants of similar size to different categories. This risk turned out to be negligible: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages was relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships between size variables plotted using data from the Biomass And Allometry Database \citep{Falster:2015}, these thresholds also align relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries. However, the groups ``juvenile'' and ``sapling'' were largely overlapping, causing us to merge these two.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``juvenile'', ``sapling'', or ``adult''.  We preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition. Using the above cut-offs we were also able to assign a stage category to studies for which no stage have been distinctly mentioned by the authors (5\% of the data, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 5 of 112 studies were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. The others equivocal cases were caused by an ambiguity in the definition of ``adult'' used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive (see ID 150). Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{fig:figA5} model.1. In the main text, we report only results using the standardized categories.

@@

We tested how robust are our results were to the particular decisions made in the analysis, in the criteria used for including a study in the analysis. The number of species used to perform the correlation, the experimental conditions, and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Thus we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the ``ideal'' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for SM (i.e. seed stage), or Hmax (i.e. adult stage).
We also tested the sensitivity of our results to inclusion of the few studies that measured a large number of responses for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and then re-calculated the effect size. This produced a reduced dataset, where multiple responses for any given trait were condensed into a study-level average.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4}). The correlation between growth and SLA shifted from strongly positive at juvenile stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($\bar{r}= -0.59 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages.

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$, $n=18$). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). The growth form, the experiment types and the growth measurement were not markedly best predictors of the data (results shown in the Fig. \ref{fig:figA5}).

The results reported above were confined to correlations across species, and under non-stressed conditions ($n = 176$). A ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used values of WD from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 7 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

Overall our study met all the quality criteria required for meta-analysis in ecology (Table \ref{tab:meta}). We showed that the impact of publication bias is probably trivial in our meta-analyses (Appendix D, Fig. \ref{fig:figA7}). The extent of the heterogeneity was high in our meta-analysis,  but substantially explained by the plant stage effect for SLA and SM, as expected (Appendix D). Finally, there was no indication of a temporal change in the effect size being reported within stages (Appendix D, Fig. \ref{fig:figA9}).

@@

%TODO please insert refs below
The meta-analysis reported here demonstrates that for several key traits the influence of traits on growth rates changes as plants grow from small to large. While this point has been made recently in smaller scale studies (refs??), it is the first time that mechanistic hypotheses are confirmed across a large number of studies already available. Meta-analysis allowed us to generalize and thus to reach a consensus about the effect of traits on growth. Growth rates are known to change with size through ontogeny \citep{Condit:1993hd, Clark:1999ed, Herault:2011dd}, \textit{i.e. plants follow a growth trajectory}; species traits should be understood as influencing this trajectory and potentially re-ranking growth rates with size. This is a shift in view from the current approach in plant strategy schemes, whereby traits are thought to render species fast-growing or slow-growing throughout life.

@@

For seedlings our meta-analyses showed, as expected, a negative relationship between SM and relative growth rate (in mass or height or volume) for small plants, that dissipated as plants grew larger. This negative relationship arises because seedlings from larger seeds are begin life larger themselves, as has been extensively discussed in the literature \citep[reviewed by][]{Turnbull:2012ew}. moreover, the prediction applies only to RGR not to AGR (Appendix Fig. \ref{fig:figA8} d).

For WD the meta-analysis showed negative correlation with growth across all plant stages, as previously reported from individual studies \citep{Wright:2010tp,Ruger:2012jv}. It should be noted however that there are few tests of this proposition at seedling stage. In our model (Appendix \ref{sec:growth}) cheaper stem construction is responsible for the predicted negative correlation. Lower xylem conductance and associated lower photosynthetic potential have also been suggested \citep[reviewed by][]{Chave:2009iy}.

As far as we are aware, our study is the first to compared correlations between Aarea and growth among small and large plants. In the growth model (Appendix \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. In juvenile stage, where most of the data were, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

Light and water micro-environments can affect growth rates, and might modify correlations with traits in relation to plant size. For example, instead of a positive relationship that disappeared with plant size, \citet{Iida:2014ep} found SLA was negatively correlated with RGR at small size and positively correlated with RGR at larger sizes. Similarly, instead of a negative correlation between WD and growth across all plant stages, \citep{Iida:2014ep,Iida:2014hq} found a negative correlation at small size that disappeared when trees were larger. \citet{Iida:2014hq} explained the switch in directionality by changes in light and water condition up the vertical profile within a forest: larger trees reached a position in the canopy where they were released from light limitations, other resources such as water may have become limiting, and the performance of trees may then have been regulated by different traits. Note that in \citet{Iida:2014hq}, the shift in the correlation between growth and WD occurred for trees larger than 22-26 cm dbh. This points to one of the possible limitations in our meta-analysis, where trees with height above 5 m were pooled as ``adult''.

@@

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Martìnez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to Westoby and a discovery grant to Falster.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig1.pdf}
\caption{Number of recorded correlations between trait and growth rate in our dataset, byt size and trait. Results shown here are for the ``complete'' dataset (recorded both across and within species, under unstressed and stressed conditions). Traits are: Aarea (or Amass) = rate of CO$_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax = asymptotic maximum height,  Ks = stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea (or NARmass) = net mass assimilation rate per unit of leaf area (or leaf mass), Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, thickness = leaf thickness, Vessel size = stem conduit size, vessel density = stem conduit density, WD = wood density.}
\label{fig:fig1}
\end{figure}

>>>>

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Anais Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia\\
\textasteriskcentered Email for correspondence: \texttt{anais.gibert@gmail.com}\\
Word count: ~XXXX words}
\date{}

@@

\item Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits (specific leaf area, wood density, maximum height, seed mass and assimilation rate) using a meta-analysis of 112 studies ($>$ 500 correlations). As a first approximation of plants size, we analysed the shift in the correlations between growth and traits across three plant stages (i.e. juvenile, sapling and adult).

@@

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Wright:2004jb,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, a growing number of empirical studies indicates the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with high photosynthetic rates and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation is mostly much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The growing discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

@@

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between traits and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a first approximation of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

High WD increases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table \ref{tab:trade-offs}), which should result in slower growth. Although expensive wood decreases the risk of mortality, this benefit is not reflected in growth outcomes. Thus increasing WD is predicted to decrease growth across the entire size range.

@@

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed in either absolute (AGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and either SLA or SM in seedlings reverts if absolute AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). For SLA, the correlation with RGR$_\textrm{mass}$ is positive but becomes negative with AGR$_\textrm{mass}$. For SM the correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For the others traits (WD, Hmax and Aarea), the predicted relationships are not affected by the growth measurement used.

@@

We aimed to identify studies that measured the correlation between five prominent functional traits (SLA, WD, Aarea, Hmax and SM) and a growth measurement across species. Our literature search was conducted in two phases.

@@

While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are limited to meta-data (i.e. title, keywords and abstracts, \citealt{Beckmann:2012hn}). This limitation does not apply to Google Scholar, where all text that is electronically available on the Internet will be indexed. Thus, we used Google Scholar to perform a more systematic search of the literature for SLA and WD.

@@

Overall, we recorded coefficient correlations for a list of 19 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 14 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. SM and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 19 traits were captured by a search on SLA and WD.

@@

We were unable to obtain a consistent and continuous measure of plant size across all studies, thus we relied on categorical groupings by ``stage'' to test our hypotheses. The authors of the 112 articles used in the meta-analyses, reported their plants as belonging to any of fix different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, by compiling data across multiple studies and vegetation types, we were concerned about the risk of assigning plants of similar size to different categories. This risk turned out to be negligible: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages was relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships between size variables plotted using data from the Biomass And Allometry Database \citep{Falster:2015}, these thresholds also align relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries. However, the group ``juvenile'' was largely overlapping with the three others, causing us to merge this category with one of the other on a case by case basis for each study.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``juvenile'', ``sapling'', or ``adult''.  We preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition. Using the above cut-offs we were also able to assign a stage category to studies for which no stage have been distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 112 articles were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. Others equivocal cases were caused by an ambiguity in the definition of ``adult'' used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive \citep{King:2006dg,King:2006he}. Whereas plants with a small stature -because stunnded by frequent wind for example \citep{Stratton:2001ck}- are assessed as adult because they are reproductive. Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{fig:figA6} model.1. In the main text, we report only results using the standardized categories.

@@

We tested how robust our results were to the particular decisions made in the analysis.
In partiuclar, we tested the sensitivity of our results to inclusion of studies that measured a large number of responses for the same trait. The idea is to check that our results are not generated by the dominance of one or few studies reporting several values for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and then re-calculated the effect size. This produced a reduced dataset, where multiple responses for any given trait were condensed into a study-level average.
We also tested if the criteria used for including a study in the analysis affect our results. In addition of the chosen criteria, the experimental conditions (stress vs unstressed), and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Thus we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the ''ideal'' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for SM (i.e. seed stage), or Hmax (i.e. adult stage).

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between growth and SLA shifted from strongly positive at juvenile stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages. The relationship between growth and either SLA or SM in seedlings is expected to differ if absolute GR in mass is used instead of RGR mass. Here, for both traits growth in mass was exclusively expressed in relative terms for seedlings(results shown in the Fig. \ref{fig:figA8} and Fig \ref{fig:figA4} f-j)

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). The growth form, the experiment types and the growth measurement were not markedly best predictors of the data (results shown in the Fig. \ref{fig:figA6}).

The results reported above were confined to correlations across species, and under non-stressed conditions ($n = 176$). A ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used values of WD from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

Overall our study met all the quality criteria required for meta-analysis in ecology (Table \ref{tab:meta}). We showed that the impact of publication bias is probably trivial in our meta-analyses (Appendix E, Fig. \ref{fig:figA7}). The extent of the heterogeneity was high in our meta-analysis,  but substantially explained by the plant stage effect for SLA and SM, as expected (Appendix E). Finally, there was no indication of a temporal change in the effect size being reported within stages (Appendix E, Fig. \ref{fig:figA9}).

@@

The meta-analysis reported here demonstrates that for several key traits the influence of traits on growth rates changes as plants grow from small to large. While the possibility has been raised recently that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}, it is the first time that mechanistic hypotheses are corroborated across a large number of studies already available. Meta-analysis allowed us to generalize and thus to reach a consensus about the effect of traits on growth. Growth rates are known to change with size through ontogeny \citep{Condit:1993hd, Clark:1999ed, Herault:2011dd}, \textit{i.e. plants follow a growth trajectory}; species traits should be understood as influencing this trajectory and potentially re-ranking growth rates with size. This is a shift in view from the current approach in plant strategy schemes, whereby traits are thought to render species fast-growing or slow-growing throughout life.

@@

For seedlings our meta-analyses showed, as expected, a negative relationship between SM and relative growth rate (in mass or height or volume) for small plants, that dissipated as plants grew larger. This negative relationship arises because seedlings from larger seeds are begin life larger themselves, as has been extensively discussed in the literature \citep[reviewed by][]{Turnbull:2012ew}.

For WD the meta-analysis showed negative correlation with growth across all plant stages, as previously reported from individual studies \citep{Wright:2010tp,Ruger:2012jv}. It should be noted however that there are few data available to test this proposition at seedling stage (n=2-5). In our model (Appendix \ref{sec:growth}) cheaper stem construction is responsible for the predicted negative correlation. Lower xylem conductance and associated lower photosynthetic potential have also been suggested \citep[reviewed by][]{Chave:2009iy}.

As far as we are aware, our study is the first to compared correlations between Aarea and growth among small and large plants. In the growth model (Appendix \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. Meta-analyses showed a positive correlation for adult and juvenile but not for saplings (n=2) In juvenile stage, where most of the data were, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

Light and water micro-environments can affect growth rates, and might modify correlations with traits in relation to plant size. For example, instead of a positive relationship that disappeared with plant size, \citet{Iida:2014ep} found SLA was negatively correlated with RGR at small size and positively correlated with RGR at larger sizes. Similarly, instead of a negative correlation between WD and growth across all plant stages, \citep{Iida:2014ep,Iida:2014hq} found a negative correlation at small size that disappeared when trees were larger. \citet{Iida:2014hq} explained the switch in directionality by changes in light and water condition up the vertical profile within a forest: larger trees reached a position in the canopy where they were released from light limitations, other resources such as water may have become limiting, and the performance of trees may then have been regulated by different traits. Note that in \citet{Iida:2014hq}, the shift in the correlation between growth and WD occurred for trees larger than 22-26 cm dbh. This points to one of the possible limitations in our meta-analysis, where trees with diameter above 10 cm were pooled as ``adult''.

@@

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Martìnez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to MW and a discovery grant to DF.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig1.pdf}
\caption{Number of recorded correlations between trait and growth rate in our dataset, by size and trait. Traits are: Aarea (or Amass) = rate of CO$_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax = asymptotic maximum height,  Ks = stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea = net mass assimilation rate per unit of leaf area, Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, thickness = leaf thickness, Vessel size = stem conduit size, vessel density = stem conduit density, WD = wood density. Results shown here are the ``complete'' dataset (see methods for details).}
\label{fig:fig1}
\end{figure}
